---
output:
  word_document: default
  html_document: default
---
# BAN 502 - Model Validation Assignment

## Khayrayyah Haamid-Day

```{r message=TRUE, warning=TRUE, include=FALSE}
library(tidyverse)
library(lubridate)
library(tidymodels)

bike <- read_csv("bike_cleaned.csv")
bike = bike %>% mutate(dteday = mdy(dteday)) %>% 
  mutate_if(is.character, as.factor) %>%
  mutate(hr = as.factor(hr)) 
```

**Task 1**

```{r}
set.seed(1234)
bike_split = initial_split(bike, prop = 0.70, strata = count)
train = training(bike_split)
test = testing(bike_split)
```

**Task 2**

There are 12,163 rows in training and 5,216 in testing.

```{r}
bike_recipe = recipe(count ~ season + mnth + hr + holiday + weekday + temp + weathersit, train) %>%
  step_dummy(all_nominal_predictors())

bike_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(bike_model) %>% 
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, train)

summary(lm_fit$fit$fit$fit)
```
This model has a good r-squared with several variables being significant though there are some variables that have a negative coefficient when they should be positive so there is some multicollinearity occuring. 

**Task 4**

```{r}
predict_train = predict(lm_fit,train)
head(predict_train)

ggplot(predict_train,aes(.pred))+
  geom_histogram()
```
There is a normal distribution in this histogram.

**Task 5** 

```{r}
lm_fit %>% predict(test) %>% bind_cols(test) %>% metrics(truth = count, estimate = .pred)
```

The R-Squared value on the testing set is very close to the R-Squared from the training model.
