 # BAN 502 - Course Project Phase 2
 
 ## Khayrayyah Haamid-Day
 
```{r include=FALSE}
library(tidymodels)
library(tidyverse)
library(rpart)
library(rpart.plot)
library(rattle)
library(caret)
library(vip)
ames_student <- read_csv("ames_student.csv")
```
 
```{r}
ames_new = ames_student %>% select(Above_Median,Lot_Area,Neighborhood,Garage_Type,House_Style,Fireplace_Qu,Year_Remod_Add,Overall_Qual,Year_Built,Overall_Cond,BsmtFin_Type_1,Full_Bath,Exter_Qual,Foundation,Bsmt_Qual,Total_Bsmt_SF,Mo_Sold,Kitchen_Qual) %>%
  mutate_if(is.character, as.factor) %>%
  mutate(Above_Median = fct_recode(Above_Median, "No" = "0", "Yes" = "1"))

summary(ames_new)
```
 
```{r}
set.seed(1234) 
ames_split = initial_split(ames_new, prop = 0.7, strata = Above_Median) #70% in training
train = training(ames_split)
test = testing(ames_split)
```
 
 
```{r}
logistic_model = 
  logistic_reg() %>% 
  set_engine("glm") 

logistic_recipe = recipe(Above_Median ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())  

logreg_wf = workflow() %>%
  add_recipe(logistic_recipe) %>% 
  add_model(logistic_model)

logistic_fit = fit(logreg_wf, train)

summary(logistic_fit$fit$fit$fit)
```
Significant - Lot_Area (positive), Full_Bath (positive), Total_Bsmt_SF (positive),Detached Garage (negative)
Some Significance - Garage_Type_BuiltIn (positive) Kitchen_Qual_Fair (negative), Foundation_PConc (positve), BsmtFin_Type_1_Unf (negative), Overall_Cond_Good (positive), Overall_Qual_Very_Good (positive), Overall_Qual_Good (positive), Overall_Qual_Average (negative), House_Style_Two_Story (positive), House_Style_One_Story (negative)

```{r}
#Second Log Model
logistic_model = 
  logistic_reg() %>% 
  set_engine("glm") 

logistic_recipe = recipe(Above_Median ~ Lot_Area + Full_Bath + Total_Bsmt_SF + Garage_Type + BsmtFin_Type_1 + Overall_Qual + Overall_Cond + Foundation, train) %>%
  step_dummy(all_nominal(), -all_outcomes())   

logreg_wf = workflow() %>%
  add_recipe(logistic_recipe) %>% 
  add_model(logistic_model)

logistic_fit2 = fit(logreg_wf, train)

summary(logistic_fit2$fit$fit$fit)
```
Compared to the first model; positives and negatives remain fairly consistant but there are more variables that are significant that were not in previous model.

```{r}
classif_recipe = recipe(Above_Median  ~., train)

tree_model = decision_tree() %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>%
  set_engine("rpart", model = TRUE) %>% #don't forget the model = TRUE flag
  set_mode("classification")

classif_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(classif_recipe)

classif_fit = fit(classif_wflow, train)

tree = classif_fit %>% 
  extract_fit_parsnip() %>% 
  pluck("fit")

fancyRpartPlot(tree)
```

```{r}
classif_fit$fit$fit$fit$cptable

treepred = predict(classif_fit, train, type = "class")
head(treepred)

confusionMatrix(treepred$.pred_class,train$Above_Median,positive="Yes")

treepred_test = predict(classif_fit, test, type = "class")
head(treepred_test)

confusionMatrix(treepred_test$.pred_class,test$Above_Median,positive="Yes")
```

```{r}
random_recipe = recipe(Above_Median ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest() %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

random_wkflw = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(random_recipe)

random_fit = fit(random_wkflw, train)
```

```{r}
predRF = predict(random_fit, train)
head(predRF)

confusionMatrix(predRF$.pred_class, train$Above_Median, positive = "Yes")

testpredrf = predict(random_fit, test)
head(testpredrf)

confusionMatrix(testpredrf$.pred_class, test$Above_Median, positive = "Yes")
```

```{r}
random_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```


