---
output:
  word_document: default
  html_document: default
---
# BAN 502 - Random Forest

### Khayrayyah Haamid-Day

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra)
library(vip)
library(ranger)
library(skimr)
drug <- read_csv("drug_data-1.csv")
```
```{r}
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")

#str(drug)

drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"

drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels =
                              c("Under16", "At16", "At17", "At18", "SomeCollege",
                                "ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
  mutate(Country = factor(Country, labels = c("USA", "NewZealand", "Other", "Australia",
                                             "Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity, labels = c("Black", "Asian", "White", "White/Black", "Other",
                                                  "White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
  select(-ID)
#str(drug_clean)

drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)
```

```{r}
#Task 1
#skim(drug_clean)

#Task 2

set.seed(1234)
drugclean_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine)
train = training(drugclean_split)
test = testing(drugclean_split)

#str(drug_clean)
```

```{r}
#Task 3
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill") 
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill") 
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill") +
  scale_x_discrete(guide = guide_axis(n.dodge=2))
grid.arrange(p1,p2,p3)
```

This shows that younger people and females are more likely to use nicotine. In terms of education, it looks like less education increases liklihood of using nicotine.

```{r}
p5 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p4,p5)
```

Most countries have higher amounts of nicotine use except for New Zealand where there is significantly lower nicotine use. Asian and black ethnicities have less nicotine use compared to the other ethnicities.

```{r}
p1 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```

These scores do not appear to have a significant impact on nicotine use. There are some outliers. 

```{r}
p5 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p6 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p7 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()
(p5)
(p6)
(p7)
```

It appears there is a relationship with nicotine use, SS and Impulsiveness. 

```{r}
#Task 4 
set.seed(123)
rf_folds = vfold_cv(train, v = 5)

nicotine_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% 
  set_engine("ranger", importance = "permutation") %>%
  set_mode("classification")

nicotine_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(nicotine_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  nicotine_wflow,
  resamples = rf_folds,
  grid = rf_grid 
)
```
```{r}
#Visualization
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

```{r}
#Task 5 
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  nicotine_wflow,
  best_rf
)

final_rf
final_rf_fit = fit(final_rf, train)

final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

The top most important variables appear to be sensation seeing, openness to experiences, being from the UK, and an age range of 45-54

```{r}
#Task 6
trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```
```{r}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

The accuracy drops a lot from the training set being 91% to 69% in the testing set, indicating a overfit model. 

**Task 7 **

This model could be used to predict a persons likelihood of using nicotine and as a result, target those with increased risk. I would recommend this model for real-world use but I would be concerned with the overfitting so I believe there would need to be some tweaking.