# Simple Linear Regression Assignment Quiz

```{r}
library(tidyverse)
library(tidymodels)
library(GGally)
library(lmtest)
library(esquisse)
```

```{r}
air = airquality

str(air)
summary(air)
ggpairs(air)
```

There are 6 variables and 153 observations and there is missing data. The Y variable is likely to be Ozone.

```{r}
air2 = drop_na(air)
```

This new data frame has 111 observations and 6 variables

```{r}
ggpairs(air2)
ggcorr(air2,label = "TRUE")
```

The most strongly correlated with Ozone is Temp and the least correlated is Day

```{r}
ggplot(air2,aes(Temp,Ozone)) +
  geom_point() +
  geom_smooth(method = "lm",se=FALSE)

esquisser()
```

This appears to be a linear relationship

```{r}
air2model = recipe(Ozone ~ Temp, air2)

lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(air2model)

lm_fit = fit(lm_wflow, air2)
summary(lm_fit$fit$fit$fit)
confint(lm_fit$fit$fit$fit)
```

a. The predictor variable is significant and the r-squared is okay

b. Likely falls within 67.3 to 71.5??

```{r}
testdata = data.frame(Temp = c(80))
predict(lm_fit, new_data = testdata)
```

```{r}
dwtest(lm_fit$fit$fit$fit)

air2 = air2 %>% mutate(resid1 = lm_fit$fit$fit$fit$residuals) #add the model residuals to our data frame
ggplot(air2,aes(x=Temp,y=resid1)) + geom_point() + theme_bw()

ggplot(air2,aes(x=resid1)) + geom_histogram() + theme_bw()
```

1. There is no evidence of non-independent residuals

2. There isn't a linear relationship

3.Data is not normally distributed and is skewed to the left